---
title: "Table 2: TEAE by SOC/PT"

format:
  html:

    theme:
      light: [litera, styles.scss]
      dark:  [darkly, styles.scss]

    toc: true
    toc-depth: 3
    number-sections: true
    smooth-scroll: true
    anchor-sections: true
    code-tools: true
    code-fold: true
    code-overflow: wrap
    df-print: paged
    fig-align: center
    fig-dpi: 300
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
  cache: false
editor: visual
---

**Add this at the top to ensure dependencies are installed during rendering.**

```{r}
options(repos = c(CRAN = "https://cloud.r-project.org"))
```

```{r}
# setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
```

**Install + load packages (run once; safe to re-run)**

```{r}
pacman::p_load(haven, dplyr, tidyr, purrr, glue, lubridate, stringr, EDCimport, pharmaRTF, data.table, gtsummary, gt, flextable, hms)
```

**Prepare the data**

```{r}
adsl <- read_xpt("adsl.xpt")

sbj <- bind_rows(
           adsl |>  mutate(trtn = trt01an, trt = trt01a),
           adsl |>  mutate(trtn = 99,     trt = "Overall")) |> 
           filter(saffl=="Y")


adae <- read_xpt("adae.xpt")

ae <- bind_rows(
           adae |>  mutate(trtn = trtan, trt = trta),
           adae |>  mutate(trtn = 99,     trt = "Overall")) |> 
           filter(saffl=="Y",trtemfl=="Y")
```

**Get the columns headers with big N and Treatment Name**

```{r}
bign <- sbj |>
  distinct(usubjid, trtn, trt) |>
  count(trtn, trt, name = "N")

finalsbj <- sbj |>
  left_join(bign, by = c("trtn", "trt")) |>
  mutate(trtlab = glue("{trt} (N= {N})")) 
 
#finalae <- ae |>
#  left_join(bign, by = c("trtn", "trt")) |>
#  mutate(trtlab = glue("{trt} (N= {N})"))

bigN <- finalsbj |> distinct(trtn, trtlab, N) |> 
  arrange(trtn)

bigN
```

**Impute Zeros for Treatment Arms with N = 0**

```{r}
dummytrt<- tibble(trtn = c(1, 2, 3, 99))
```

```{r}
temptrtN<-sbj |> count(trtn, trt)
temptrtN
```

```{r}
trtN<-dummytrt  |> 
    left_join(temptrtN ,by=c("trtn")) |> 
    mutate(N=if_else(is.na(n),0,n)) |> 
    select(-n)

trtN
```

**TOP row : Overall: Number of subjects with at least one TEAE.**

```{r}
sub_n<- ae |> 
    group_by(trtn) |> 
    summarise(label= "Overall", count = n_distinct(usubjid), events=n()) |> 
    ungroup()

sub_n
```

**SOC rows: Number of subjects with at least one TEAE in the specified SOC.**

```{r}
soc_n <- ae |>
    group_by(aebodsys, trtn) |>
    summarise(
    count  = n_distinct(usubjid),
    events = n(),
    .groups = "drop"        # or "drop_last" / "keep"
  )

soc_n
```

**PT Rows: Number of subjects with at least one TEAE in the specified PT.**

```{r}
pt_n <- ae |>
    group_by(aebodsys, aedecod,trtn) |>
    summarise(
    count  = n_distinct(usubjid),
    events = n(),
    .groups = "drop"
  )

pt_n
```

**Combine Top Row (Overall), SOC-level, and PT-level counts into a single dataset and replace missing values (NAs) with blanks.**

```{r}
all_Ntemp <- bind_rows(sub_n, soc_n, pt_n) |> 
    mutate(across(c(label, aebodsys, aedecod), ~ if_else(is.na(.), "", .)))
```

**Create zero counts for treatment groups where an event is not observed.**

-   Get all available SOC and PT values (Overall, SOC, and SOC/PT combinations).

-   Create one row per treatment group for each Overall/SOC/PT row.

```{r}
dummy_ae<-all_Ntemp |> 
    complete(nesting(label,aebodsys,aedecod),trtn=c(1,2,3,99)) |> 
    select(-count,-events)
dummy_ae
```

**Merge the dummy AE grid with the observed counts.\
For rows not present in the data, set subject counts and event counts to 0.**

```{r}
all_Ntemp2<-dummy_ae |> 
    left_join(all_Ntemp,by=c("label","aebodsys", "aedecod", "trtn")) |>
    mutate(across(c(count,events),~if_else(is.na(.),0,.)))

all_Ntemp2
```

Merge with the denominator Big N count for percentage calculation

```{r}
all_N<-all_Ntemp2 %>%
    left_join(bigN,by="trtn")
```

**Compute the percentage column and concatenate with count**

```{r}
all_Perc<-all_N |> 
    mutate(
    percent=if_else(N!=0,count/N*100,0),
    percentc=str_c(" (",sprintf("%.1f",percent),"%)"),
    cp=if_else(count==0,"0",str_c(count, percentc))
 )

all_Perc
```

**Create the row labels column;**

```{r}
prefinal1<-all_Perc |> 
    mutate(
    label=case_when(
    aebodsys=="" & aedecod=="" ~ label,
    aebodsys!="" & aedecod=="" ~ aebodsys,
    aebodsys!="" & aedecod!="" ~ str_c("\u00A0\u00A0\u00A0",aedecod),
    TRUE~""
 )
 )

prefinal1
```

```{r}
prefinal2 <- prefinal1 |> 
    pivot_wider(
    id_cols = c(aebodsys, aedecod, label),
    names_from = trtlab,
    values_from = cp) |> 
 arrange(aebodsys,aedecod,label)
 
prefinal2
```

```{r}
# view(counts06)
final <- prefinal2 |> 
    select( -aebodsys, -aedecod )

glimpse(final)
```

```{r}
tlf_gt <- function(df,
                   title = "Table X. Treatment-Emergent Adverse Events by System Organ Class and Preferred Term",
                   subtitle = "Safety Population",
                   stub_width = gt::px(420),   # <-- add
                   trt_width  = gt::px(90)) {  # <-- optional

  df <- as.data.frame(df)

  stub     <- names(df)[1]
  trt_cols <- names(df)[-1]

  # Section header rows = all treatment cells blank/NA
  is_section <- apply(df[, trt_cols, drop = FALSE], 1, function(x) {
    all(is.na(x) | trimws(as.character(x)) == "")
  })

  # Sub-rows = stub starts with spaces
  is_sub <- grepl("^\\s+", df[[stub]])
  df[[stub]] <- sub("^\\s+", "", df[[stub]])   # remove spaces; we’ll indent via gt

  # 2-line column headers: "Placebo (N= 19)" -> "Placebo<br>(N=19)"
  lab_trt <- setNames(lapply(trt_cols, function(x) {
    x2 <- gsub("\\s+", " ", x)
    gt::html(sub("\\s*\\(N\\s*=\\s*([0-9]+)\\s*\\)\\s*$",
                 "<br>(N=\\1)", x2, perl = TRUE))
  }), trt_cols)

  labs <- c(setNames(list(gt::html("")), stub), lab_trt)

  # ---- key fix: embed width values in formulas (no 'stub_width' symbol to resolve later) ----
  f_stub <- rlang::new_formula(1, stub_width, env = environment())                   # col 1
  f_trt  <- rlang::new_formula(dplyr::all_of(trt_cols), trt_width, env = environment())

  g <- gt::gt(df) %>%
    gt::tab_header(
      title = gt::md(paste0("**", title, "**")),
      subtitle = subtitle
    ) %>%
    gt::opt_row_striping() %>%
    gt::cols_width(f_stub, f_trt) %>%                      # <-- widen first col
    gt::cols_align("left",   columns = 1) %>%              # <-- use col position (robust)
    gt::cols_align("center", columns = dplyr::all_of(trt_cols)) %>%
    gt::opt_table_font(font = list("Courier New", "Consolas", "monospace")) %>%
    gt::tab_options(
      table.font.size = gt::px(12),
      data_row.padding = gt::px(2),

      table.border.top.style = "solid",
      table.border.top.width = gt::px(2),
      column_labels.border.bottom.style = "solid",
      column_labels.border.bottom.width = gt::px(2),
      table.border.bottom.style = "solid",
      table.border.bottom.width = gt::px(2),

      table_body.hlines.style = "none",
      table_body.vlines.style = "none",
      column_labels.vlines.style = "none"
    )

  # apply labels (programmatically)
  g <- do.call(gt::cols_label, c(list(g), labs))

  # Bold section headers (first column)
  g <- g %>%
    gt::tab_style(
      style = gt::cell_text(weight = "bold"),
      locations = gt::cells_body(columns = 1, rows = is_section)
    ) %>%
    # Indent sub-rows (first column)
    gt::tab_style(
      style = gt::cell_text(indent = gt::px(18)),
      locations = gt::cells_body(columns = 1, rows = is_sub & !is_section)
    ) %>%
    gt::tab_source_note(gt::md("*Percentages are based on the column N.*"))

  g
}


```

```{r}

# tlf_gt(final, stub_width = gt::px(420))
# or
# tlf_gt(final, stub_width = gt::pct(60))

tlf_gt(final, stub_width = gt::px(420), trt_width = gt::px(85))


```
